\chapter{Dynamic Programming}


\section{Palindromic Subsequence}

\subsection{Longest Palindromic Subseqence}

\sssec{Problem Statment:}

Given a sequence, find the length of its Longest Palindromic Subsequence(LPS). 
In a palindromic subsequence, elements
read the same backward and forward.

A subsequence is a sequence that can be derived from another sequence by deleting some 
or no elements without changing the order of the remaining elements.

\sssec{Examples:}

    \textit{Input:} "abdbca"
    
    \textit{Output:} 5
    
    \textit{Explanation:} LPS is "abdba"
    \newline

    \textit{Input:} "cddpd"
    
    \textit{Output:} 3
    
    \textit{Explanation:} LPS is "ddd"

\sssec{Basic Solution:}

A basic brute-force solution could be to try all the subsequences of the given sequence.
We can start processing from the beginning and the end of the sequence. So at any step, we have two options:

\begin{enumerate}
    \item If str[begin]==str[end], increment counter by two. Subproblem: LPS in str[begin+1][end-1].
    \item If str[begin]!=str[end], nothing. Subproblem: max LPS in str[begin+1][end] and LPS in str[begin][end-1].
\end{enumerate}

\sssec{Bottom Up Idea:}


    n = string.length $\Rightarrow$ dp[n][n] and the subproblem is $dp[i][j]:$ the LPS in substring str[i:j].

    The solution to dp[i][j]:

    \begin{enumerate}
        \item str[i] == str[j] $\Rightarrow$  dp[i+1][j-1]+2
        \item str[i] != str[j] $\Rightarrow$ max(dp[i+1][j],dp[i][j-1])
    \end{enumerate}

    The solution to the LPS original problem is dp[0][n] which is the top right corner. Also, in $2\times 2$ block, top right corner
    is based on the surrounding three.

    $\Rightarrow$ populate the table in the following order:
    \begin{enumerate}
        \item From left to right.
        \item From bottom to top.
    \end{enumerate}

    \sssec{Bottom Up Code:}

\begin{lstlisting}[language=C++,basicstyle=\tiny]
    class LPS{
        public:
            int findLPSLength(const string &st){
                vector<vector<int>> dp(st.length(), vector<int>(st.length(), 0));

                // every sequence with one element is a palindrome of length 1
                for (int i = 0; i < st.length(); i++) {
                    dp[i][i] = 1;
                }

                // rows from bottom to up
                for (int startIndex = st.length() - 1; startIndex >= 0; startIndex--) {
                    
                    // column from left to right
                    for (int endIndex = startIndex + 1; endIndex < st.length(); endIndex++) {
                        
                        // case 1: elements at the beginning and the end are the same
                        if (st[startIndex] == st[endIndex]) {
                            
                            dp[startIndex][endIndex] = 2 + dp[startIndex + 1][endIndex - 1];
                        } 
                        else { // case 2: skip one element either from the beginning or the end
                            
                            dp[startIndex][endIndex] =
                                max(dp[startIndex + 1][endIndex], dp[startIndex][endIndex - 1]);
                        }
                    }         
                }

                return dp[0][st.length() - 1];
            }
        }
\end{lstlisting}

\newpage

\subsection{Longest Palindromic Substring:}

\sssec{Problem Statment}


Given a string, find the length of its Longest Palindromic Substring (LPS). 
In a palindromic string, elements read the same backward and forward.


\sssec{Examples:}

    \textit{Input:} "abdbca"
    
    \textit{Output:} 3
    
    \textit{Explanation:} LPS is "bdb"
    \newline

    \textit{Input:} "cddpd"
    
    \textit{Output:} 3
    
    \textit{Explanation:} LPS is "dpd"



